---
title: "Spatial blocks creation for train/test split"
author: "Francesco Pasanisi"
date: "October 12, 2023"
output:
  html_document:
    df_print: paged
project: "Using high-resolution satellite imagery and deep learning to map artisanal mining spatial extent in the DRC"
---

## 0. Install and import needed packages

```{r}
if(!"blockCV" %in% installed.packages()){install.packages("blockCV")}
if(!"sf" %in% installed.packages()){install.packages("sf")}

library(blockCV)
library(sf)
```

## 1. Stream data from Zenodo repository

```{r}
asm_sites_url <- "https://zenodo.org/record/8379225/files/asm_sites.geojson"
study_area_url <- "https://zenodo.org/record/8379225/files/study_area.geojson"
```

```{r}
asm_sites = st_read(asm_sites_url)
study_area = st_read(study_area_url)
```

## 2. Create spatial blocks and do the split

Here spatial blocks of 100 km of range are created. This range is controlled by the 'size' parameter of cv_spatial(). 'hexagon' is set to FALSE in order to have square blocks.

```{r}
blocks <- cv_spatial(x = asm_sites,
                 hexagon = F,
                 size = 100000,
                 k = 10)
```

Hereafter there is the train/test split. The created blocks are treated in order to have approximately the 70% of points assigned to the training set and the remaining 30% to the test set.

```{r}
block_assignments <- blocks$folds_ids
unique_blocks <- unique(block_assignments)

# 70% of training data is considered
train_blocks <- sample(unique_blocks, size = round(length(unique_blocks) * 0.7))

# make the actual split
train_data <- asm_sites[block_assignments %in% train_blocks, ]
test_data <- asm_sites[!block_assignments %in% train_blocks, ]
```

Let's visualize the split!

```{r}
spatial_blocks <- blocks$blocks # extract blocks geometry

plot(st_geometry(asm_sites), col = "grey", pch = 16, cex = 0.5)
plot(st_geometry(train_data), col = "blue", pch = 16, cex = 0.5, add = TRUE)
plot(st_geometry(test_data), col = "red", pch = 16, cex = 0.5, add = TRUE)
plot(st_geometry(spatial_blocks), col = NA, border = "black", add = T)
legend("topright", legend = c("Training Data", "Test Data"), fill = c("blue", "red"))
```

## 3. Export data

```{r}
# set the working directory here
setwd("your/working/directory")

st_write(train_data, "train_data.geojson")
st_write(test_data, "test_data.geojson")

# st_write(spatial_blocks, "spatial_blocks.geojson")
```
